---
title: "final Project"
author: "Prasannaw Maddipati"
date: "2024-04-26"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**Loading the dataset**

```{r}
df_airbnb <- read.csv("Airbnb_Data.csv")
head(df_airbnb)
```

```{r}
str(df_airbnb)
```

```{r}
library(tidyverse)

city <- ggplot(df_airbnb, aes(x = city, fill = city)) +
  geom_bar() +
  labs(x = "City", y = "Frequency", title = "Distribution of Airbnb Listings by City") +
  scale_fill_brewer(palette = "Set3")


city + geom_text(stat = "count", aes(label = ifelse(..count.. != 0, ..count.., "")), vjust = -0.5)
```

**Filtering out Airbnb dataset only for SF city.**

```{r}
df_airbnb2 <- df_airbnb[df_airbnb$city == 'SF',]
df_airbnb2 <- df_airbnb2[, c("log_price", "property_type", "room_type", "accommodates", "bathrooms", "bed_type", "cancellation_policy", "cleaning_fee", "host_has_profile_pic", "host_identity_verified", "host_response_rate", "instant_bookable", "neighbourhood", "number_of_reviews", "review_scores_rating", "zipcode", "bedrooms", "beds")]

head(df_airbnb2)
str(df_airbnb2)
```

Analysing the distribution of the variables

```{r}
library(tidyverse)
custom_palette <- scales::hue_pal()(35)

type <- ggplot(df_airbnb2, aes(x = property_type, fill = property_type)) +
  geom_bar(color = "black") +
  labs(x = "Property Type", y = "Frequency", title = "Distribution of Airbnb Listings By Property Type") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) +
  scale_fill_manual(values = custom_palette)

type + geom_text(stat = "count", aes(label = ifelse(..count.. != 0, ..count.., "")), vjust = -0.5, position = position_stack(vjust = 0.5))
```

```{r}
r <- ggplot(df_airbnb2, aes(x = factor(room_type))) +
  geom_bar(fill = "darkred", color = "black") +
  labs(x = "Room Type", y = "Frequency", title = "Distribution of Airbnb Listings by Room Type") +
  scale_x_discrete(labels = function(x) paste0(x))  


r + geom_text(stat = "count", aes(label = ifelse(..count.. != 0, ..count.., "")), vjust = -0.5)
```

```{r}
a <- ggplot(df_airbnb2, aes(x = factor(accommodates))) +
  geom_bar(fill = "darkred", color = "black") +
  labs(x = "Number of accommodates", y = "Frequency", title = "Distribution of Airbnb Listings by Number of accommodates") + scale_x_discrete(labels = function(x) paste0(x))

a + geom_text(stat = "count", aes(label = ifelse(..count.. != 0, ..count.., "")), vjust = -0.5)
```

```{r}
p <- ggplot(df_airbnb2, aes(x = factor(bathrooms))) +
  geom_bar(fill = "darkred", color = "black") +
  labs(x = "Number of Bathrooms", y = "Frequency", title = "Distribution of Airbnb Listings by Number of Bathrooms") +
  scale_x_discrete(labels = function(x) paste0(x))+ theme(plot.title = element_text(size=9))



p + geom_text(stat = "count", aes(label = ifelse(..count.. != 0, ..count.., "")), vjust = -0.5)
```

```{r}
bedroom <- ggplot(df_airbnb2, aes(x = factor(bedrooms))) +
  geom_bar(fill = "brown", color = "black") +
  labs(x = "Number of Bedrooms", y = "Frequency", title = "Distribution of Airbnb Listings by Number of Bedrooms") +
  scale_x_discrete(labels = function(x) paste0(x))  


bedroom + geom_text(stat = "count", aes(label = ifelse(..count.. != 0, ..count.., "")), vjust = -0.5)
```

```{r}
bed_type <- ggplot(df_airbnb2, aes(x = factor(bed_type))) +
  geom_bar(fill = "darkred", color = "black") +
  labs(x = "Number of bed_type", y = "Frequency", title = "Distribution of Airbnb Listings by Number of bed_type") +
  scale_x_discrete(labels = function(x) paste0(x))  


bed_type + geom_text(stat = "count", aes(label = ifelse(..count.. != 0, ..count.., "")), vjust = -0.5)
```

```{r}
beds <- ggplot(df_airbnb2, aes(x = factor(beds))) +
  geom_bar(fill = "darkred", color = "black") +
  labs(x = "Number of beds", y = "Frequency", title = "Distribution of Airbnb Listings by Number of beds") +
  scale_x_discrete(labels = function(x) paste0(x))


beds + geom_text(stat = "count", aes(label = ifelse(..count.. != 0, ..count.., "")), vjust = -0.5)
```

```{r}
fee <- ggplot(df_airbnb2, aes(x = cancellation_policy)) +
  geom_bar(fill = "darkred", color = "black") +
  labs(x = "Cancellation Policy", y = "Frequency", 
       title = "Distribution of Airbnb Cancellation Fees")

fee + geom_text(stat = "count", aes(label = ifelse(..count.. != 0, ..count.., "")), vjust = -0.5)

```

```{r}
cleaning <- ggplot(df_airbnb2, aes(x = cleaning_fee)) +
  geom_bar(fill = "darkred", color = "black") +
  labs(x = "Cleaning Fee", y = "Frequency", 
       title = "Distribution of Airbnb Cleaning Fees")

cleaning + geom_text(stat = "count", aes(label = ifelse(..count.. != 0, ..count.., "")), vjust = -0.5)
```

```{r}
verified <- ggplot(df_airbnb2, aes(x = host_identity_verified)) +
  geom_bar(fill = "darkred", color = "black") +
  labs(x = "Host Has Verified Identity", y = "Frequency", 
       title = "Distribution of Airbnb Host Verified Identity")

verified + geom_text(stat = "count", aes(label = ifelse(..count.. != 0, ..count.., "")), vjust = -0.5)
```

```{r}
Profile <- ggplot(df_airbnb2, aes(x = host_has_profile_pic)) +
  geom_bar(fill = "darkred", color = "black") +
  labs(x = "Host Has Profile pic", y = "Frequency", 
       title = "Distribution of Airbnb Host Profile Pic")

Profile + geom_text(stat = "count", aes(label = ifelse(..count.. != 0, ..count.., "")), vjust = -0.5)
```

```{r}
Response <- ggplot(df_airbnb2, aes(x = host_response_rate)) +
  geom_bar(fill = "darkred", color = "black") +
  labs(x = "Host Has Response Rate", y = "Frequency", 
       title = "Distribution of Airbnb Host Response Rate") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

Response + geom_text(stat = "count", aes(label = ifelse(..count.. != 0, ..count.., "")), vjust = -0.5)
```

```{r}
bookable <- ggplot(df_airbnb2, aes(x = instant_bookable)) +
  geom_bar(fill = "darkred", color = "black") +
  labs(x = "instant_bookable", y = "Frequency", 
       title = "Distribution of Airbnb instant_bookable") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

bookable + geom_text(stat = "count", aes(label = ifelse(..count.. != 0, ..count.., "")), vjust = -0.5)
```

```{r}
neighbourhood <- ggplot(df_airbnb2, aes(x = neighbourhood)) +
  geom_bar(fill = "darkred", color = "black") +
  labs(x = "neighbourhood", y = "Frequency", 
       title = "Distribution of Airbnb neighbourhood") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

neighbourhood + geom_text(stat = "count", aes(label = ifelse(..count.. != 0, ..count.., "")), vjust = -0.5)
```

```{r}
zipcode <- ggplot(df_airbnb2, aes(x = zipcode)) +
  geom_bar(fill = "darkred", color = "black") +
  labs(x = "zipcode", y = "Frequency", 
       title = "Distribution of Airbnb zipcode") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

zipcode + geom_text(stat = "count", aes(label = ifelse(..count.. != 0, ..count.., "")), vjust = -0.5)
```

Notice 'zipcode' variable have duplicate levels with decimal places.
Removing the duplicate values and regrouping.

```{r}
unique(df_airbnb2$zipcode)

unique(substr(df_airbnb2$zipcode, 1, 5))
```

```{r}
df_airbnb2$zipcode <- substr(df_airbnb2$zipcode, 1, 5)
zipcode <- ggplot(df_airbnb2, aes(x = zipcode)) +
  geom_bar(fill = "darkred", color = "black") +
  labs(x = "zipcode", y = "Frequency", 
       title = "Distribution of Airbnb zipcode") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

zipcode + geom_text(stat = "count", aes(label = ifelse(..count.. != 0, ..count.., "")), vjust = -0.5)
```

```{r}
ggplot(df_airbnb2, aes(x = number_of_reviews)) +
  geom_histogram(binwidth = 10, fill = "darkred", color = "black") +
  labs(x = "Number of Reviews", y = "Frequency", title = "Distribution of Number of Reviews")

```

```{r}
ggplot(df_airbnb2, aes(x = review_scores_rating)) +
  geom_histogram(binwidth = 5, fill = "darkred", color = "black") +
  labs(x = "Review Scores Rating", y = "Frequency", title = "Distribution of Review Scores Rating")
```

```{r}
ggplot(df_airbnb2, aes(x = log_price)) +
  geom_histogram( fill = "darkred", color = "black") +
  labs(x = "log Price", y = "Frequency", title = "Distribution of log prices of airbnb's")

ggplot(df_airbnb2, aes(x = exp(log_price))) +
  geom_histogram( fill = "darkred", color = "black") +
  labs(x = "Price", y = "Frequency", title = "Distribution of actual prices of airbnb's")
```

```{r}
str(df_airbnb2)
```

```{r}
#Removing null values
df_airbnb3 <- na.omit(df_airbnb2)

#Converting categorical variables into factors
df_airbnb3$property_type <- as.factor(df_airbnb3$property_type)
df_airbnb3$room_type <- as.factor(df_airbnb3$room_type)
df_airbnb3$accommodates <- as.factor(df_airbnb3$accommodates)
df_airbnb3$bathrooms <- as.factor(df_airbnb3$bathrooms)
df_airbnb3$bed_type <- as.factor(df_airbnb3$bed_type)
df_airbnb3$cancellation_policy <- as.factor(df_airbnb3$cancellation_policy)
df_airbnb3$cleaning_fee <- factor(df_airbnb3$cleaning_fee)
df_airbnb3$host_has_profile_pic <- as.factor(df_airbnb3$host_has_profile_pic)
df_airbnb3$host_identity_verified <- as.factor(df_airbnb3$host_identity_verified)
df_airbnb3$host_response_rate <- as.factor(df_airbnb3$host_response_rate)
df_airbnb3$instant_bookable <- as.factor(df_airbnb3$instant_bookable)
df_airbnb3$neighbourhood <- as.factor(df_airbnb3$neighbourhood)
df_airbnb3$bedrooms <- as.factor(df_airbnb3$bedrooms)
df_airbnb3$beds <- as.factor(df_airbnb3$beds)
df_airbnb3$zipcode <- as.factor(df_airbnb3$zipcode)

pairs(log_price ~ accommodates + bathrooms + cleaning_fee + number_of_reviews + review_scores_rating + bedrooms + beds, data = df_airbnb3)
```

```{r}

```

```{r}
str(df_airbnb3)
```

```{r}
# Fitting a multiple linear regression baseline model
lm1 <- lm(log_price ~ ., data = df_airbnb3)
summary(lm1)
```

The R-squared vale us 0.696. Let us check the model diagnostic plots.

```{r}
plot(lm1)
hist(lm1$residuals)
```

```{r}
shapiro.test(resid(lm1))
```

Since our sample size is large, let us perform Anderson-Darling
normality test.

```{r}
library(nortest)
ad.test(resid(lm1))
```

Doing box-cox transformation of the response variable

```{r}
library(MASS)
library(car)

boxcox(lm1)
summary(p1 <- powerTransform(lm1))
p1$roundlam #get the value of rounded power to transform later
```

```{r}
# Applying power transformation to the response variable along with the step function to remove unimportant variables.
lm2 <- lm(bcPower(log_price, p1$roundlam) ~ ., data = df_airbnb3)
lm_step1 <- step(lm2, trace=0)
summary(lm_step1)
```

```{r}
plot(lm_step1)
hist(lm_step1$residuals)
```

Building a model with exp(log_price) as the response variable.

```{r}
lm3 <- lm(exp(log_price) ~ ., data = df_airbnb3)
summary(lm3)
```

```{r}
boxcox(lm3)
summary(p2 <- powerTransform(lm3))
lm4 <- lm(bcPower(exp(log_price), p2$roundlam) ~ ., data = df_airbnb3)
lm_step2 <- step(lm4, trace=0)
summary(lm_step2)
```

```{r}
summary(lm1)$adj.r.squared #baseline model
summary(lm_step1)$adj.r.squared #baseline model with power transformation

summary(lm3)$adj.r.squared #exponential model
summary(lm_step2)$adj.r.squared #exponential model with power transformation
```

Upon performing the power Transformation, the adjusted R-squared value
of lm_step2 model improved when compare. So far, lm1 model has the
highest adjusted R-squared value of 0.683 however **lm_step1** model is
parsimonious than lm1 with an adjusted R-squared value of 0.68. Since
the difference among the values is minute, let us choose **lm_step1**
model and perform model diagnostics.

```{r}
# Checking the frequency distribution of each variable
my_func <- function(df, group){
 df %>%
   group_by_(group) %>%
   summarise(my_count = n()) %>%
   arrange(desc(my_count))
}

my_func(df_airbnb3, 'property_type')
my_func(df_airbnb3, 'room_type')
my_func(df_airbnb3, 'accommodates')
my_func(df_airbnb3, 'bathrooms')
my_func(df_airbnb3, 'bed_type')
my_func(df_airbnb3, 'cancellation_policy')
my_func(df_airbnb3, 'cleaning_fee')
my_func(df_airbnb3, 'host_has_profile_pic')
my_func(df_airbnb3, 'host_identity_verified')
my_func(df_airbnb3, 'host_response_rate')
my_func(df_airbnb3, 'instant_bookable')
my_func(df_airbnb3, 'neighbourhood')
my_func(df_airbnb3, 'number_of_reviews')
my_func(df_airbnb3, 'review_scores_rating')
my_func(df_airbnb3, 'zipcode')
my_func(df_airbnb3, 'bedrooms')
my_func(df_airbnb3, 'beds')
```

```{r}
summary(lm_step1)
```

```{r}
#Regrouping the variables based on the significance levels in the model summary. Checking if regrouping improves the model performance.

library(tidyverse)

df_airbnb4 <- df_airbnb3
df_airbnb4 <- mutate(df_airbnb4, bathrooms2 = case_when(
  bathrooms == "5" | bathrooms == "8"  ~ '5', TRUE ~ bathrooms))

df_airbnb4 <- mutate(df_airbnb4, property_type2 = case_when(
  property_type == "Boat" | property_type == "Castle" | 
  property_type == "Cave" | property_type == "Condominium" | 
  property_type == "Dorm" | property_type == "House" | 
  property_type == "Serviced apartment" ~ property_type, TRUE ~ 'Other'))
```

```{r}
str(df_airbnb4)
```

```{r}
df_airbnb4$bathrooms2 <- as.factor(df_airbnb4$bathrooms2)
df_airbnb4$property_type2 <- as.factor(df_airbnb4$property_type2)
df_airbnb4$accommodates2 <- as.factor(df_airbnb4$accommodates)

lm5 <- lm(log_price ~ property_type2 + room_type + accommodates2 +
            bathrooms2  + bed_type + cancellation_policy +
            host_identity_verified  + host_response_rate +
            neighbourhood  +  number_of_reviews + review_scores_rating + 
            zipcode + bedrooms + beds, data = df_airbnb4)
summary(lm5)
```

```{r}
boxcox(lm5)
summary(p3<- powerTransform(lm5))
lm6 <- lm(bcPower(log_price, p3$roundlam) ~ property_type2 + room_type + accommodates2 +
            bathrooms2  + bed_type + cancellation_policy +
            host_identity_verified  + host_response_rate +
            neighbourhood  +  number_of_reviews + review_scores_rating + 
            zipcode + bedrooms + beds, data = df_airbnb4)
summary(lm6)
```

lm5 model has the highest adjusted R squared value of 0.6845

```{r}
# checking for multicollinearity
vif(lm5)
```

```{r}
plot(lm5)
```

```{r}
set.seed(999) # set seed for reproducibility
n <- nrow(df_airbnb4)
n
floor(0.7*n)
```

```{r}
# randomly sample 70% of rows for training set
train <- sample(1:n, 3520)
# fit model using training data
lm_train <- lm(log_price ~ property_type2 + room_type + accommodates2 + 
    bathrooms2 + bed_type + cancellation_policy + host_identity_verified + 
    host_response_rate + neighbourhood + number_of_reviews + 
    review_scores_rating + zipcode + bedrooms + beds, data=df_airbnb4,
subset = train)

summary(lm_train)
```

```{r}
test <- df_airbnb4[-train, ]
preds_test <- predict(lm_train, newdata = test)
tb <- table(prediction = preds_test,
actual <- df_airbnb4$log_price)
addmargins(tb)
```

Since property_type2 of Serviced apartment is not present in the test
data set, we are getting the above error. To avoid this issue, we have
to ensure both training and testing data sets have similar distribution.

We can skip model building for now as it is out of scope for STAT632
project.
